---
title: "Homework 1 - Time Series Visualization"
author: "Prof. Dr. Stephan Trahasch"
date: 'Submission date: 17.12.2020'
output:
  html_document:
    theme: cerulean
    css: styles/homework.css
    toc: yes
    toc_depth: 3
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
options(digits = 5)

# libraries to load
library(fpp3)
library(fpp2) 
```

# Exercise 1
Use the help menu to explore what the series `gold`, `woolyrnq` and `gas` represent. These are available in the `forecast` package.

```{r, echo=TRUE, cache=TRUE, eval=FALSE}
help(gold)
help(woolyrnq)
help(gas)
# to do

```

Use `autoplot` to plot each of these in separate plots.

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(gold)
autoplot(woolyrnq)
autoplot(gas)
```

**What is the frequency of each commodity series?**
Hint: apply the `frequency()` function.

```{r, echo=TRUE, cache=TRUE}
# to do
frequency(gold)
frequency(woolyrnq)
frequency(gas)
```

Use `which.max()` to spot the outlier in the `gold` series. 
**Which observation was it?**

```{r, echo=TRUE, cache=TRUE}
# to do
which.max(gold)
```

# Exercise 2 

Import the file `tute1.csv` from the folder data and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

You can read the data into R with the following script:

```{r, echo=TRUE, cache=TRUE}
tute1 <- read.csv("data/tute1.csv", header = TRUE)
View(tute1)
```

We convert the data to time series. The `[,-1]` removes the first column which contains the quarters as we don't need them now.

```{r, echo=TRUE, cache=TRUE}
mytimeseries <- ts(tute1[, -1], start = 1981, frequency = 4)
```

Construct time series plots of each of the three series and use 'facets = TRUE' .

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(mytimeseries,facets=TRUE)
```

Check what happens when you don't include `facets=TRUE`.

```{r, echo=TRUE, cache=TRUE, dependson="exmyts"}
# to do
autoplot(mytimeseries)
```

# Exercise 3

Import some monthly Australian retail data from the foler data. These represent retail sales in various categories for different Australian states, and are stored in a MS-Excel file.

You can read the data into R with the following script:

```{r, echo=TRUE, cache=TRUE}
retaildata <- readxl::read_excel("data/retail.xlsx", skip = 1)
```

The second argument (`skip=1`) is required because the Excel sheet has two header rows.

Select one of the time series as follows (but replace the column name with your own chosen column):

```{r, echo=TRUE, cache=TRUE}
myts <- ts(retaildata[, "A3349873A"], frequency =12, start = c(1990, 1))
myts
```

Explore your chosen retail time series using the following functions: `autoplot`, `ggseasonplot`, `ggsubseriesplot`, `gglagplot`, `ggAcf`

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(myts)
ggsubseriesplot(myts)
ggseasonplot(myts)
gglagplot(myts)
ggAcf(myts)
```

**Can you spot any seasonality, cyclicity and trend? What do you learn about the series?**

Your answer: Der Anstieg beginnt im seit neustem im Juli und kehrt im Oktober zur normalität zurück.Dies ist im Seasonplot zu erkennen. Generell ist ein wachstum über die Jahre festzustellen. Der Wachstumstrend erkennt man im Autoplot, hierzu kann die Trendlinie angezeigt werden. Die cyclicity kann auch anhand der Seasondaten und des Subseriesplot erkannt werden. Nach mitte des Jahres ist ein Anstieg zu verzeichnen, der nach 3 Monaten abflacht auf den ungefähr alten Wert.


# Exercise 4
Create time plots of the following time series: goog, writing, fancy, a10, h02.

    + Use `help()` to find out about the data in each series.
    + For the `goog` plot, modify the axis labels and title.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# to do
help(goog)
help(writing)
help(fancy)
help(a10)
help(h02)
autoplot(a10) + ylab("$ million") + xlab("Year") +
ggtitle("Antidiabetic drug sales")
autoplot(goog) + ylab("stock prices") + xlab("trading days") +
ggtitle("Daily closing stock prices")

```

Use the `ggseasonplot` and `ggsubseriesplot` functions to explore the seasonal patterns in the following time series:  `writing`, `fancy`, `a10`, `h02`.

- What can you say about the seasonal patterns?
- Can you identify any unusual years?


```{r, echo=TRUE}
# writing
# to do
ggseasonplot(writing)
ggsubseriesplot(writing)
```

Your answer:Im Sommer erreichen die Werte jedes Jahr ihr Tiefpunkt. Generell ist ein Anstieg in den letzten Jahren zu erkennen.


```{r, echo=TRUE}
# fancy
# to do
ggseasonplot(fancy)
ggsubseriesplot(fancy)
```

Your answer:Anfang des Jahres sind Werte auf dem Tiefpunkt, während diese am Ende des Jahres ihr Maximum erreichen.Jedes Jahr starten die Werte etwas weiter oben.


```{r, echo=TRUE}
# a10
# to do
ggseasonplot(a10)
ggsubseriesplot(a10)
```

Your answer:Die TS endet im Juni 2008. Es ist ein Wachstum der einzelnen Jahre zu erkennen. Ab 2004 wird eine Seasonalität deutlich, der maximale Wert ist am Anfang des Jahres erreicht. Der Tiefpunkt im Feburar und wächst bis zum Ende des Jahres wieder an. I den letzten Jahren gibt es starke Schwankungen.


```{r, echo=TRUE}
# h02
# to do
ggseasonplot(h02)
ggsubseriesplot(h02)
```

Your answer:Maximal Werte sind im Januar und Dezember erreicht und der Tiefpunkt im Februar. Von Februar an ist ein Wachstum bis in den Dezember zu erkennen. Die TS beginnt im Juli 1991 und ender im Juni 2008. Die letzten Jaren haben starke Schwankungen.


# Exercise 5
Use the following graphics functions: `autoplot`, `ggseasonplot`, `ggsubseriesplot`, `gglagplot`, `ggAcf` and explore features from the following time series: `hsales`, `usdeaths`, `bricksq`, `sunspotarea`, `gasoline`.

    + Can you spot any seasonality, cyclicity and trend?
    + What do you learn about the series?

Example for hsales:

```{r, echo=TRUE, cache=TRUE}
autoplot(hsales)
ggseasonplot(hsales)
ggsubseriesplot(hsales)
gglagplot(hsales)
ggAcf(hsales)
```

  + Seasonality evident in all plots
  + Cyclicity seen in first two plots
  + No trend
  + ACF only shows seasonality. Cycle length too long to show up here.

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(usdeaths)
ggseasonplot(usdeaths)
ggsubseriesplot(usdeaths)
gglagplot(usdeaths)
ggAcf(usdeaths)
```

Your answer:
  + Seasonality ist in allen Plots zu erkennen
  + Cyclicity wird in den ersten Beiden festgestellt
  + kein Trend


```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(bricksq)
ggseasonplot(bricksq)
ggsubseriesplot(bricksq)
gglagplot(bricksq)
ggAcf(bricksq)
```

Your answer:
  + Seasonality besonders im dritten Block sieht man, das Q1 am schwächsten ist.
  + Cyclicity wird in der ersten Grafik deutlicher Sichtbar
  + Trend erkennbar in Plot1, steigend bis 1975 dann eher Konstant.

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(sunspotarea)
#ggseasonplot(sunspotarea)
#ggsubseriesplot(sunspotarea)
gglagplot(sunspotarea)
ggAcf(sunspotarea)
```

Your answer:
  + nicht Seasonal
  + Cyclicity über mehrere Jahre zu erkennen
  + kein Trend
```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(gasoline)
ggseasonplot(gasoline)
#ggsubseriesplot(gasoline)
gglagplot(gasoline)
ggAcf(gasoline)
```

Your answer:
  + Seasonality in den Wintermonaten zu erkennen(Tiefpunkte)
  + keine konstante Cyclic zu erkennen
  + Trend erkennbar in Plot1, steigend bis 2007

# Exercise 6

The `pigs` data shows the monthly total number of pigs slaughtered in Victoria, Australia, from Jan 1980 to Aug 1995. Use `mypigs <- window(pigs, start=1990)` to select the data starting from 1990. Use `autoplot` and `ggAcf` for `mypigs` series and compare these to white noise plots.

```{r, echo=TRUE, cache=TRUE}
# White noise
set.seed(1711)
y <- ts(rnorm(50))
autoplot(y) + ggtitle("White noise")
```

```{r}
ggAcf(y)
```


```{r, echo=TRUE, cache=TRUE}
# todo
rawdat<-ts(pigs,frequency=12,start=c(1990,1))
mypigs <- window(pigs, start=1990)
autoplot(rawdat)
ggAcf(rawdat)
autoplot(mypigs) +
xlab("Year") + ylab("thousands") +
ggtitle("Number of pigs slaughtered in Victoria")

ggAcf(mypigs)
```

Your answer: Es scheint keine starkes Rauschen in den Daten zu geben. Es konnten keine Muster zwischen den Whitenoise und PigSlautherd festgestellte werden. Die ACF ist unter 0.23.


# Exercise 7

`dj` contains 292 consecutive trading days of the Dow Jones Index. Use `ddj <- diff(dj)` to compute the daily changes in the index. Plot `ddj` and its ACF. Do the changes in the Dow Jones Index look like white noise?

```{r}
# todo
ddj <- diff(dj)
autoplot(ddj)
ggAcf(ddj)
```

Your answer: Auch hier ist nur schwer etwas zu interpretieren, die Daten sehen nach Whitenoise aus. Die Grenze wird nur einmal überschritten. Der ACF ist bei maximal 0.12.

